Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-27 analysis/paper/paper.Rmd
trying to understand the text on CV equality testing in results section

```{r first-two-PCs-byplot, fig.cap= "Pottery shape distribution by each occupation phase according to the first two PCs"}
# the pots groups by phase, and their shape distribution according to 
# the first two principle components of the EFA 
# png(here::here("analysis/figures/pca-biplot.png"), res = 300, w = 1500, h = 1500)
# plot(op.p, 2, 
#      morpho = TRUE, 
#      ellipses = TRUE,
#      cex = 1, 
#      cex.labelsgroups = 1.5)
# dev.off()

@@

@@

To test for differences in the distributions of shape variables indicated by the PC scores and biplot, we use the multivariate analysis of variance (MANOVA) test to compare pairwise combinations across our three occupation phases. Table \@ref(tab:MANOVA-table) shows significant differences in shape between Pre-European and Post-European phases (p = $`r pval_post_pre`$), and Pre-European and Chinese phases (p = $`r pval_chi_pre`$). Although there is considerable overlap of shape variables between the Pre-European and Post-European phases, their PC scores are significantly different from each other. There is little difference between the shapes in the Post-European and Chinese contact periods.

@@

Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%) before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%).

>>>>

```{r first-two-PCs-byplot, fig.cap= "Pottery shape distribution by each occupation phase according to the first two PCs"}

@@

To test for differences in the distributions of shape variables indicated by the PC scores shown in Figure \ref(@fig:first-two-PCs-byplot)), we used a multivariate analysis of variance (MANOVA) test to compare pairwise combinations across our three occupation phases. Table \@ref(tab:MANOVA-table) shows significant differences in shape between Pre-European and Post-European phases (p = $`r pval_post_pre`$), and Pre-European and Chinese phases (p = $`r pval_chi_pre`$). Although there is considerable overlap of shape variables between the Pre-European and Post-European phases, their PC scores are significantly different from each other. There is little difference between vessel shapes in the Post-European and Chinese contact periods.

@@

@@

Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%)  before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). <!-- these numeric values in this paragraph should come from inline code, not copy-pasted in like this -->

<!-- You need to put the code that computes the MSLRT results right in here, in the Rmd -->
